Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints\thanks{This research was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2020-02-01 2019/NUMTA_Barkalov/SoftComputing/article.tex
Add address, fix tables and theorems.

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{lineno,hyperref,amsmath}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

@@

%\authorrunning{Short form of author list} % if too long for running head

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}

@@

\begin{abstract}
%Insert your abstract here. Include keywords, PACS and mathematical
%subject classification numbers as needed.
This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition. The algorithms for solving the problems of this class require the use of adequate estimates of the a priori unknown Lipschitz constants for the problem functions. A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and an underestimated one. The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy. The considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately. The convergence conditions of the proposed algorithm are formulated in the corresponding theorem. The results of the numerical experiments on a series of multiextremal problems with non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
\keywords{Global optimization \and Multiextremal problems \and Non-convex constraints \and Lipschitz constant estimates \and Dimensionality reduction \and Numerical methods}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

	The main part of the paper has the following structure. In Section 2, the index method of accounting for the constraints and the dimensionality reduction scheme for the constrained global optimization problems are described. In Section 3, a description of the computational rules of the index global search algorithm is given and its theoretical properties are formulated. In Section 4, the index algorithm with dual Lipschitz constant estimates is formulated and its convergence is proved. Section 5 presents the results of the numerical experiments carried out on several series of the multiextremal problems with non-convex constraints. Section 6 concludes the paper.

\section{Index Method of Accounting for the Constraints and Dimensionality Reduction}
\label{sec:1}
	We consider the global optimization problems of the form
\begin{equation}\label{problem}
	\min{\left\{\varphi(y):y \in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
	D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\end{equation}

@@

	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
\begin{align}\label{epsilon_reserved_solution}
	\varphi(y_{\epsilon}) &= \nonumber \\
	&\min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
\end{align}
and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
\begin{align}\label{Y_epsilon}
	Y_{\epsilon} &= \nonumber \\
	&\left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
\end{align}
Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.

\begin{figure}%[h]
	\center{\includegraphics[width=0.6\linewidth]{epsilon_reserved_solution}}
	\caption{$\varepsilon$-reserved solution.}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{align}\label{one_dimensional_problem}
	\varphi(y(x^\ast)) =  
	\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{align}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval $[0,1]$ for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

\section{Index-Based Global Search Algorithm}
\label{sec:2}
	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}). The set of triplets 
$\{x^k,\; z^k=g_{\nu}\big(y(x^k)\big),\; \nu=\nu(x^k)\},\; 1 \leq k \leq n,$
makes up the search information accumulated by the method after performing $n$ steps. The operation of the index-based global search algorithm is defined by the following rules.

@@

	The convergence conditions for the index global search algorithm can be formulated in the form of the following theorem from \cite{Strongin2000}. 
\\
\textbf{Theorem 1}. Assume that the following conditions are satisfied:
\begin{enumerate} 
  \item Problem (\ref{problem}) has an $\varepsilon$-reserved solution $y_{\varepsilon}$ from (\ref{epsilon_reserved_solution}).
  \item Functions $g_j(y), 1 \leq j \leq m+1,$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
$$
	g_j \left( y(x) \right) = G_j \left( y(x) \right), \; x \in Q_j, \; 1 \leq j \leq m+1,
$$
where the sets $Q_j$ are from (\ref{Q_intervals}), and $y(x)$ is the curve from (\ref{one_dimensional_problem}).
	\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m,$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
	\item For a sufficiently large iteration number $k$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
		\begin{equation}\label{theorem_inequalities}
			r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1. 
		\end{equation}
Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions	
\end{enumerate}
\begin{eqnarray*}
	\varphi(\bar y) =	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\ \leq \varphi(y_{\epsilon}).
\end{eqnarray*}
\textbf{Note 1}. The computational scheme of the algorithm is described in the form convenient for  formulating its theoretical properties. The software implementation of this scheme may be made more frugal. For example, according to Step 7, the point $x^{k+1}$ of the next trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e. $x^{k+1} \in (x_{t-1}, x_t)$. Therefore, the set of trial points $x^1,\ldots, x^k$ (and the related trial results (\ref{trial_result})) may not be ordered  at every step of the algorithm but may be maintained in an ordered state using appropriate data structures.
\\
\textbf{Note 2}. The use of the set of parameters $\epsilon_1,\ldots,\epsilon_m$ (the values of which correspond to the coordinates of the reserve vector $\epsilon_R$) in the index algorithm will significantly reduce the density of the trial sequence in the vicinity of the feasible domain boundaries. However, this raises a question about the particular values of the reserve vector components $\epsilon_R$. In \cite{Strongin2000}, it was proposed to choose the components of the vector $\epsilon_R$ according to the formula 
\begin{equation}\label{epsilon_nu}
	\epsilon_{\nu} = \mu_{\nu}\delta, \; 1 \leq \nu \leq m, 
\end{equation}
where $\mu_{\nu}$ are the current estimates from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter. This approach to choosing the reserve vector components allows one to take into account the properties of different problem functions characterized by their Lipschitz constants. This method for choosing the reserve vector was used in the numerical experiments, the results of which are presented below.
\\
\textbf{Note 3}. The index algorithm for solving the constrained problems is based on the information-statistical algorithm of global search (AGS) for solving unconstrained problems. AGS has been compared repeatedly with both deterministic and nature-inspired global optimization algorithms. The results of the experimental comparison (see, for example, \cite{Sovrasov2019}) show that AGS is superior to many known methods of similar purpose, including DIRECT and DIRECT\textit{l}.

@@

\section{Index Algorithm with Dual Lipschitz Constant Estimates}
\label{sec:3}
	The index global search algorithm described in the previous section (hereafter referred to as IA) is intended for solving  the multiextremal problems, in which the objective function and the constraints satisfy the Lipschitz condition.  In this case, a priori estimates of the constant values are not required for the algorithm to work. These estimates are computed in the course of global search, based on the available search information.

	According to Theorem 1, the sequence of trial points $\{y^k\}$ will converge to a point $\bar y \in Y_\epsilon$ if the condition (\ref{theorem_inequalities}) is satisfied. So far, an appropriate choice of the parameters $r_{\nu}$ allows using the values
\begin{equation}\label{estimates_Lipschitz_constants}
	\frac{r_{\nu}\mu_{\nu}}{2^{3-\frac{1}{N}}\sqrt{N+3}}, \; 1 \leq \nu \leq m+1, 
\end{equation}
as the estimates of the Lipschitz constants for the problem functions $g_{j}(y), 1 \leq j \leq m+1$.

@@

	Note that Theorem 1 on the convergence of the index method implies the generation of an infinite sequence of trials $\{y^k\}=\big\{y(x^k )\big\}$ by the algorithm. In this case, at the final stage of the search (i.e. in the case when one of the points of this trial sequence falls into the vicinity of the $\varepsilon$-reserved solution $y_\epsilon$) there always exists an interval, at one of the boundaries of which the current minimum value of the objective function $z_{m+1}^\ast$  from (\ref{z_estimates}) is achieved. It is this interval that will have the highest characteristic since it is the most promising one for performing further trials.

@@

	The first four steps of the algorithm repeat Step 1 -- Step 4 of the initial index algorithm from Section 3.

@@

	Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}$, $x_t$ are not the same, i.e.
$$
	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
$$
Otherwise, execute a trial at the point
\begin{eqnarray*}
	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \\
	\nu=\nu(x_{t-1})=\nu(x_t).
\end{eqnarray*}
When computing the point $x^{k+1}$, use the value of the parameter $r_{\nu}$ fixed at Step~6.

@@

	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
\begin{align*}
	\varphi(y_1, y_2)&=-1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
	&-\left(0.5(y_1-1)(y_2-1)\right)^4 \\
	& \exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
\end{align*}
in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
\begin{eqnarray}
	g_1(y_1, y_2) &=& 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \nonumber \\
	g_2(y_1, y_2) &=& 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \nonumber \\
	g_3(y_1, y_2) &=& 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0. \nonumber
\end{eqnarray}
The feasible domain of this problem is non-connected and consists of three subdomains. The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).

@@

	In Fig. \ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked. The best point is marked by red. For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ from (\ref{epsilon_nu}). For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used. The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$. Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used. To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method. The results of larger-scale numerical experiments are presented in Section 4.

	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
\textbf{Theorem 2}. Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled. Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1,\; 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e. satisfies the condition 
\begin{align}\label{phi_bar_y}
	\varphi(\bar y) & = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \nonumber \\
	& \leq \varphi(y_{\epsilon}).
\end{align}

\textbf{Proof}. Because for the IA with the parameters $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled, then any limit point $\bar{y}=y(\bar{x})$ of the trial sequence generated by the IA for the problem (\ref{problem}) satisfies the conditions (\ref{phi_bar_y}).

@@

\section{Conclusion}
\label{sec:5}
	In this paper, a method for solving the Lipschitz optimization problems, which constitute an important class of the global optimization problems, was considered. In applied global optimization problems, the functions are usually defined as the “black boxes”. Consequently, the values of the Lipschitz constant for them are unknown a priori, which makes their estimation one of the key problems when constructing the Lipschitz optimization methods.

>>>>

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{lineno,hyperref,amsmath,array}
\usepackage{multirow}
\usepackage[round]{natbib} 
\usepackage[misc]{ifsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints\thanks{This research was supported by the Russian Science Foundation, project No. 16-11-10150.}
}

@@

\institute{Roman Strongin        \and
					 Konstantin Barkalov (\Letter) \and
					 Semen Bevzuk \at
              Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, 603950, Russia \\
              \email{konstantin.barkalov@itmm.unn.ru}
}

@@

\begin{abstract}
This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition. The algorithms for solving the problems of this class require the use of adequate estimates of the a priori unknown Lipschitz constants for the problem functions. A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and an underestimated one. The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy. The considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately. The convergence conditions of the proposed algorithm are formulated in the corresponding theorem. The results of the numerical experiments on a series of multiextremal problems with non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
\keywords{Global optimization \and Multiextremal problems \and Non-convex constraints \and Lipschitz constant estimates \and Dimensionality reduction \and Numerical methods}
\subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

	The main part of the paper has the following structure. In Section \ref{sec:2}, the index method of accounting for the constraints and the dimensionality reduction scheme for the constrained global optimization problems are described. In Section \ref{sec:3}, a description of the computational rules of the index global search algorithm is given and its theoretical properties are formulated. In Section \ref{sec:4}, the index algorithm with dual Lipschitz constant estimates is formulated and its convergence is proved. Section \ref{sec:5} presents the results of the numerical experiments carried out on several series of the multiextremal problems with non-convex constraints. Section \ref{sec:6} concludes the paper.

\section{Index Method of Accounting for the Constraints and Dimensionality Reduction}
\label{sec:2}
	We consider the global optimization problems of the form
\begin{equation}\label{problem}
	\min{\left\{\varphi(y):y \in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
	D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\end{equation}

@@

	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
\begin{align}\label{epsilon_reserved_solution}
	\varphi(y_{\epsilon}) & = \nonumber \\
	& \min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
\end{align}
and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
\begin{align}\label{Y_epsilon}
	Y_{\epsilon} & = \nonumber \\
	& \left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
\end{align}
Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.

\begin{figure}
	\center{\includegraphics[width=0.6\linewidth]{epsilon_reserved_solution}}
	\caption{$\varepsilon$-reserved solution.}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{align}\label{one_dimensional_problem}
	& \varphi(y(x^\ast)) =  \nonumber \\
	& \min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{align}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

\section{Index-Based Global Search Algorithm}
\label{sec:3}
	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}). The set of triplets 
$\{x^k,\; z^k=g_{\nu}\big(y(x^k)\big),\; \nu=\nu(x^k)\},\; 1 \leq k \leq n,$
makes up the search information accumulated by the method after performing $n$ steps. The operation of the index-based global search algorithm is defined by the following rules.

@@

	The convergence conditions for the index global search algorithm can be formulated in the form of the following theorem from \cite{Strongin2000}. 
\\
\begin{theorem}\label{theorem:1} Assume that the following conditions are satisfied:
	\begin{enumerate} 
		\item Problem (\ref{problem}) has an $\varepsilon$-reserved solution $y_{\varepsilon}$ from (\ref{epsilon_reserved_solution}).
		\item Functions $g_j(y), 1 \leq j \leq m+1,$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
	$$
		g_j \left( y(x) \right) = G_j \left( y(x) \right), \; x \in Q_j, \; 1 \leq j \leq m+1,
	$$
	where the sets $Q_j$ are from (\ref{Q_intervals}), and $y(x)$ is the curve from (\ref{one_dimensional_problem}).
		\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m,$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
		\item For a sufficiently large iteration number $k$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
			\begin{equation}\label{theorem_inequalities}
				r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1. 
			\end{equation}
	\end{enumerate}
	Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions	
	\begin{eqnarray*}
		\varphi(\bar y) =	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\
		\leq \varphi(y_{\epsilon}).
	\end{eqnarray*}
\end{theorem}
\textbf{Note 1}. The computational scheme of the algorithm is described in the form convenient for  formulating its theoretical properties. The software implementation of this scheme may be made more frugal. For example, according to Step 7, the point $x^{k+1}$ of the next trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e. $x^{k+1} \in (x_{t-1}, x_t)$. Therefore, the set of trial points $x^1,\ldots, x^k$ (and the related trial results (\ref{trial_result})) may not be ordered  at every step of the algorithm but may be maintained in an ordered state using appropriate data structures.
\\
\textbf{Note 2}. The use of the set of parameters $\epsilon_1,\ldots,\epsilon_m$ (the values of which correspond to the coordinates of the reserve vector $\epsilon_R$) in the index algorithm will significantly reduce the density of the trial sequence in the vicinity of the feasible domain boundaries. However, this raises a question about the particular values of the reserve vector components $\epsilon_R$. In \cite{Strongin2000}, it was proposed to choose the components of the vector $\epsilon_R$ according to the formula 
\begin{equation}\label{epsilon_nu}
	\epsilon_{\nu} = \mu_{\nu}\delta, \; 1 \leq \nu \leq m, 
\end{equation}
where $\mu_{\nu}$ are the current estimates from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter. This approach to choosing the reserve vector components allows one to take into account the properties of different problem functions characterized by their Lipschitz constants. This method for choosing the reserve vector was used in the numerical experiments, the results of which are presented below.
\\
\textbf{Note 3}. The index algorithm for solving the constrained problems is based on the information-statistical algorithm of global search (AGS) for solving unconstrained problems. AGS has been compared repeatedly with both deterministic and nature-inspired global optimization algorithms. The results of the experimental comparison (see, for example, \cite{Sovrasov2019}) show that AGS is superior to many known methods of similar purpose, including DIRECT and DIRECT\textit{l}.

@@

\section{Index Algorithm with Dual Lipschitz Constant Estimates}
\label{sec:4}
	The index global search algorithm described in the previous section (hereafter referred to as IA) is intended for solving  the multiextremal problems, in which the objective function and the constraints satisfy the Lipschitz condition.  In this case, a priori estimates of the constant values are not required for the algorithm to work. These estimates are computed in the course of global search, based on the available search information.

	According to Theorem \ref{theorem:1}, the sequence of trial points $\{y^k\}$ will converge to a point $\bar y \in Y_\epsilon$ if the condition (\ref{theorem_inequalities}) is satisfied. So far, an appropriate choice of the parameters $r_{\nu}$ allows using the values
\begin{equation}\label{estimates_Lipschitz_constants}
	\frac{r_{\nu}\mu_{\nu}}{2^{3-\frac{1}{N}}\sqrt{N+3}}, \; 1 \leq \nu \leq m+1, 
\end{equation}
as the estimates of the Lipschitz constants for the problem functions $g_{j}(y), 1 \leq j \leq m+1$.

@@

	Note that Theorem \ref{theorem:1} on the convergence of the index method implies the generation of an infinite sequence of trials $\{y^k\}=\big\{y(x^k )\big\}$ by the algorithm. In this case, at the final stage of the search (i.e. in the case when one of the points of this trial sequence falls into the vicinity of the $\varepsilon$-reserved solution $y_\epsilon$) there always exists an interval, at one of the boundaries of which the current minimum value of the objective function $z_{m+1}^\ast$  from (\ref{z_estimates}) is achieved. It is this interval that will have the highest characteristic since it is the most promising one for performing further trials.

@@

	The first four steps of the algorithm repeat Step 1 -- Step 4 of the initial index algorithm from Section \ref{sec:3}.

@@

	Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}$, $x_t$ are not the same, i.e.
$$
	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
$$
Otherwise, execute a trial at the point
\begin{eqnarray*}
	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \\
	\nu = \nu(x_{t-1})=\nu(x_t).
\end{eqnarray*}
When computing the point $x^{k+1}$, use the value of the parameter $r_{\nu}$ fixed at Step~6.

@@

	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
\begin{align*}
	\varphi(y_1, y_2) & = -1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
	& -\left(0.5(y_1-1)(y_2-1)\right)^4 \\
	& \exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
\end{align*}
in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
\begin{eqnarray*}
	g_1(y_1, y_2) & = & 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \\
	g_2(y_1, y_2) & = & 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \\
	g_3(y_1, y_2) & = & 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0.
\end{eqnarray*}
The feasible domain of this problem is non-connected and consists of three subdomains. The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point \\ $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).

@@

	In Fig. \ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked. The best point is marked by red. For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ from (\ref{epsilon_nu}). For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used. The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$. Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used. To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method. The results of larger-scale numerical experiments are presented in Section \ref{sec:4}.

	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
\begin{theorem}\label{theorem:2} Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem \ref{theorem:1} are fulfilled. Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1,\; 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e. satisfies the condition 
\begin{align}\label{phi_bar_y}
	\varphi(\bar y) & = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \nonumber \\
	& \leq \varphi(y_{\epsilon}).
\end{align}
\end{theorem}

\textbf{Proof}. Because for the IA with the parameters $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem \ref{theorem:1} are fulfilled, then any limit point $\bar{y}=y(\bar{x})$ of the trial sequence generated by the IA for the problem (\ref{problem}) satisfies the conditions (\ref{phi_bar_y}).

@@

\section{Conclusion}
\label{sec:6}
	In this paper, a method for solving the Lipschitz optimization problems, which constitute an important class of the global optimization problems, was considered. In applied global optimization problems, the functions are usually defined as the “black boxes”. Consequently, the values of the Lipschitz constant for them are unknown a priori, which makes their estimation one of the key problems when constructing the Lipschitz optimization methods.
